---
title: "Penalized Regression in R"
author: "Priyansh Desai"
date: "2023-03-21"
output: html_document
---

Loading necessary libraries

```{r}
library(caret)
library(tidyverse)
library(glmnet)
library(caTools)
```

Loading the Dataset and splitting into training and testing set 

```{r}
questionMark.data <- read.csv('QuestionMark.csv')
cat('There are', sum(is.na(questionMark.data)))
data <- na.omit(questionMark.data)
cat("There are", nrow(data), " left in the dataset.")
```

```{r}
set.seed(123)
training.samples <- data$y %>%
  createDataPartition(p = 0.75, list = FALSE)
train.data <- data[training.samples, ]
test.data <- data[-training.samples, ]
dim(train.data)  # 1097 observations 
dim(test.data)  # 363 observations  
```

RIDGE REGRESSION

```{r}
x <- model.matrix(y~. , train.data)[,-1]
y <- train.data$y
cv <- cv.glmnet(x, y, alpha = 0)
cv$lambda.min
```

The best lambda for Ridge Regression is 129.0571. 

```{r}
model <- glmnet(x, y, alpha = 0, lambda = cv$lambda.min)
coef(model)
```

Making prediction for checking RMSE and Coefficient of Determination Rsquare.

```{r}
x.test <- model.matrix(y~. , test.data)[,-1]
predictions <- model %>% predict(x.test) %>% as.vector()
data.frame(
  RMSE = RMSE(predictions, test.data$y),
  Rsquare = R2(predictions, test.data$y)
)
```
RMSE for Ridge Regression is 9009.3678 and the Coefficient of Determination is 0.6729274.

```{r}
plot(y = predictions, x = test.data$y, xlab = 'Observed data', ylab = 'Estimated data')
abline(1,1)
```

LASSO REGRESSION

```{r}
cv <- cv.glmnet(x, y, alpha = 0.1)
cv$lambda.min
```

The best lambda for LASSO is 10.96667

```{r}
model <- glmnet(x, y, alpha = 1, lambda = cv$lambda.min)
coef(model)
```
```{r}
x.test <- model.matrix(y~. , test.data)[,-1]
predictions <- model %>% predict(x.test) %>% as.vector()
data.frame(
  RMSE = RMSE(predictions, test.data$y),
  Rsquare = R2(predictions, test.data$y)
)
```
The RMSE for LASSO is 958.4058 and the Coefficient of Determination is 0.6571835.

```{r}
plot(y = predictions, x = test.data$y, xlab = 'Observed data', ylab = 'Estimated data')
abline(1,1)
```

ELASTIC NET MODEL

```{r}
model <- train(y~. , data = train.data, method = "glmnet", trControl = trainControl("cv", number = 10),
               tuneLength = 10)
model$bestTune
```

Alpha = 1 and lambda = 7.351186.

```{r}
coef(model$finalModel, model$bestTune$lambda)
```

```{r}
model <- glmnet(x, y, alpha = 1, lambda = 7.351186)
x.test <- model.matrix(y~. , test.data)[,-1]
predictions <- model %>% predict(x.test)
data.frame(
  RMSE = RMSE(predictions, test.data$y),
  Rsquare = R2(predictions, test.data$y)
)
```

The RMSE = 964.2876 and the Coefficient of Determination = 0.6560664.

```{r}
plot(y = predictions, x = test.data$y, xlab = "Observed data", ylab = "Estimated Data")
abline(1,1)
```

