---
title: "ams580_hw3"
author: "Mustafa Yigit Isik"
date: "2023-02-20"
output: word_document
---
#**Load Packages**
```{r}
if (!requireNamespace("caTools")) install.packages('caTools')
if (!requireNamespace("tidyverse")) install.packages('tidyverse')
if (!requireNamespace("caret")) install.packages('caret')
if (!requireNamespace("rpart")) install.packages('rpart')
if (!requireNamespace("rattle")) install.packages('rattle')
library(caTools)
library(tidyverse)
library(caret)
library(rpart)
library(rattle)
```

#__Read the data__
```{r}
data <- read.csv("/Users/mustafayigitisik/Desktop/stuff/semesters/spring 2023/ams 580/hw3_CART/banknote.csv")

#removes potential rows where at least one cell is empty 
data <- na.omit(data)
cat('There are', nrow(data), 'rows left.')
data$class <- as.factor(data$class)
```

#**Split Data**
```{r}
set.seed(123)
training.samples <- data$class %>% 
  createDataPartition(p = 0.8, list = FALSE)
training  <- data[training.samples, ]
testing <- data[-training.samples, ]
```

#**Full Tree and its Confusion Matrix ~ Problem 2**
```{r}
model <- rpart(class ~., data = training, control = rpart.control(cp=0))
par(xpd = NA)
fancyRpartPlot(model)
```
```{r}
pred <- predict(model,newdata = testing, type ='class')
pred <- ifelse(pred == 1, 'real', 'forged')
table(pred,testing$class)
```
#**Prune the Data ~ Problem 3**

To make tree more robust, weprune the full tree using the training data with 10-fold cross-validation. Please (1) show the complexity plot, (2) report the best CP value, and (3) draw the pruned tree using rattle.
```{r}
set.seed(123)
model2 <- train(
  class ~., data = training, method = "rpart",
  trControl = trainControl("cv", number = 10),
  tuneLength = 100)
plot(model2)
```
```{r}
model2$bestTune
```
```{r}
fancyRpartPlot(model2$finalModel)
```

#**Check testing data via pruned version ~ Problem 4**

Using optimal pruned tree to predict whether each check is forged or not in the testing data.

Compute the confusion matrix and report the sensitivity, specificity and the overall accuracy for the testing data.
```{r}
pred <- predict(model2, newdata = testing)
pred <- ifelse(pred == 1, 'real' , 'forged')
matrix <- table(pred,testing$class)
matrix
```
Sensitivity
```{r}
matrix[2,2]/(matrix[1,2] + matrix[2,2])
```
Specificity
```{r}
matrix[1,1]/(matrix[1,1] + matrix[2,1])
```
Accuracy
```{r}
(matrix[1,1]+matrix[2,2])/(sum(matrix))
```

#**Regression Model ~ Problem 5**
```{r}
data$class <- as.factor(data$class)
model3 <- glm(class ~ . , data = training, family = binomial)
summary(model3)
```

#Confusion Matrix based on Reg
```{r}
predict_training<-predict(model3, training, type= 'response' )
predict_testing<-predict(model3, testing, type= 'response' )
confusion_matrix_training <- table(actual = training$class, prediction = predict_training > 0.8)
confusion_matrix_testing <- table(actual = testing$class, prediction = predict_testing > 0.8)
confusion_matrix_training
```

#Accuracy
```{r}
accuracy_training<-sum(diag(confusion_matrix_training))/sum(confusion_matrix_training)
accuracy_training
accuracy_testing<-sum(diag(confusion_matrix_testing))/sum(confusion_matrix_testing)
accuracy_testing
```
#Sensitivity
```{r}
sensitivity_training <-confusion_matrix_training[2,2]/(confusion_matrix_training[2,2]+confusion_matrix_training[2,1])
sensitivity_training
sensitivity_testing <-confusion_matrix_testing[2,2]/(confusion_matrix_testing[2,2]+confusion_matrix_testing[2,1])
sensitivity_testing
```
#Specificity
```{r}
specificity_training = confusion_matrix_training[1,1]/(confusion_matrix_training[1,1]+confusion_matrix_training[1,2])
specificity_training
specificity_testing = confusion_matrix_testing[1,1]/(confusion_matrix_testing[1,1]+confusion_matrix_testing[1,2])
specificity_testing
```
#Problem 6

**From majority voting classifier, we import our findings [115, 5, 10, 147] here to test for sensitivity, accuracy, and specificity**
```{r}
majority <- c(115, 5, 10, 147)
majority_matrix <- matrix(majority, nrow=2)
majority_matrix
```

Sensitivity
```{r}
majority_matrix[2,2]/(majority_matrix[1,2] + majority_matrix[2,2])
```
Specificity
```{r}
majority_matrix[1,1]/(majority_matrix[1,1] + majority_matrix[2,1])
```
Accuracy
```{r}
(majority_matrix[1,1]+majority_matrix[2,2])/sum(majority_matrix)
```