---
title: "ams580_quiz3"
output: word_document
author: Mustafa Yigit Isik
date: "2023-02-22"
---
#**Load Packages**
```{r}
if (!requireNamespace("caTools")) install.packages('caTools')
if (!requireNamespace("tidyverse")) install.packages('tidyverse')
if (!requireNamespace("caret")) install.packages('caret')
if (!requireNamespace("rpart")) install.packages('rpart')
if (!requireNamespace("rattle")) install.packages('rattle')
if (!requireNamespace("xlsx")) install.packages('xlsx')

library(caTools)
library(tidyverse)
library(caret)
library(rpart)
library(rattle)
library(xlsx)
```

#__Read the data__
```{r}
data <- read.csv("/Users/mustafayigitisik/Desktop/stuff/semesters/spring 2023/ams 580/quizzes/quiz 3/GreatUnknown.csv")

#removes potential rows where at least one cell is empty 
data <- subset(data, !is.na(w1) | !is.na(w2) | !is.na(w3) | !is.na(w4) | !is.na(w5) | !is.na(w6) | !is.na(w7) | !is.na(w8) | !is.na(w9) | !is.na(w10)
| !is.na(w11)| !is.na(w12)| !is.na(y))
cat('There are', nrow(data), 'rows left.')
data$y <- as.factor(data$y)
```

#**Split Data**
```{r}
set.seed(456)
training.samples <- data$y %>% 
  createDataPartition(p = 0.75, list = FALSE)
training  <- data[training.samples, ]
testing <- data[-training.samples, ]
```

#**Full Tree, its Confusion Matrix, and Sens/Spec/Accuracy ~ Problem 2**
```{r}
model <- rpart(y ~., data = training, control = rpart.control(cp=0))
par(xpd = NA)
fancyRpartPlot(model)
```

```{r}
pred1 <- predict(model,newdata = testing, type ='class')
pred1 <- ifelse(pred1 == 1, '1', '0')
confusionMatrix(factor(pred1), factor(testing$y), positive = '1')
```

#**Prune the Data ~ Problem 3**

To make tree more robust, we prune the full tree using the training data with 10-fold cross-validation. Please (1) show the complexity plot, (2) report the best CP value, and (3) draw the pruned tree using rattle.
```{r}
# set.seed(456)
model2 <- train(
  y ~., data = training, method = "rpart",
  trControl = trainControl("cv", number = 10),
  tuneLength = 100)
plot(model2)
```
```{r}
model2$bestTune
```
```{r}
fancyRpartPlot(model2$finalModel)
```

#**Check testing data via pruned version ~ Problem 4**

Using optimal pruned tree to predict whether each check is 0 or 1 in the testing data.

Compute the confusion matrix and report the sensitivity, specificity and the overall accuracy for the testing data.
```{r}
pred2 <- predict(model2, newdata = testing)
pred2 <- ifelse(pred2 == 1, '1' , '0')
confusionMatrix(factor(pred2), factor(testing$y), positive = '1')
```

#**Regression Model ~ Problem 5**
```{r}
data$y <- as.factor(data$y)
model3 <- glm(y ~ . , data = training, family = binomial)
summary(model3)
```

#Confusion Matrix based on Reg
```{r}
probabilities <- model3 %>% predict(testing, type = "response")
pred3 <- ifelse(probabilities > 0.5, 1, 0)
confusionMatrix(factor(pred3), factor(testing$y), positive = '1')
```
#**Problem 6**

```{r}
update.test = cbind(testing, pred1, pred2, pred3)
write.xlsx(update.test, 'testingquiz3.xlsx', sheetName = 'testingquiz3')

pred = cbind(pred1, pred2, pred3)
pred.m = apply(pred,1,function(x) names(which.max(table(x)))) 
pred.m = as.numeric(pred.m)
confusionMatrix(factor(pred.m), factor(testing$y), positive = '1')
```